---
title: "Practical Machine Learning Project"
author: "EAA"
output: html_document
date: '2022-05-30'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview of the data set and methodology

 

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website [here:](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har)) (see the section on the Weight Lifting Exercise Dataset)


## Load Required Packages
```{r load packages, message = FALSE}
library(knitr)
library(tidyverse)
library(caret)
library(randomForest)
library(rpart)
```

## Loading and Processing Data


The training data for this project can be found [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv).

The test data is available [here.](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)

```{r process}
  
train_set <- read_csv("pml-training.csv", na = c("", "#DIV/0!", "NA"), show_col_types = FALSE)
test_set <- read_csv("pml-testing.csv", na = c("", "#DIV/0!", "NA"), show_col_types = FALSE)

```
##  Cleaning and Tidying Data
```{r clean}
 
## Data cleaning
# Delete entire columns with only NAs

#train_set <- train_set[colSums(!is.na(train_set)) > 0 ]


## Delete entire columns with only NAs

train_set <- train_set[, colSums(is.na(train_set)) == 0]
test_set <- test_set[, colSums(is.na(test_set)) == 0]

## Delete irrelevant columns (or variables) not used for prediction


train_set <- select(train_set, -c(1:7))

```

# Cross-Validation
```{r, cross}

## Split the train_set into two: trainData and testData

inTrain <- createDataPartition(y = train_set$classe, list = FALSE, p = 0.7)
train_data <- train_set[inTrain, ]
test_data <- train_set[-inTrain, ]

```

# Prediction

```{r, model}

# Building Decision Tree Model

#Decision Tree

# Fit model on train_data

dtree_modfit <- rpart(classe ~., data = train_data, method = "class")

# use model to predict classe in validation set (test_data)

dtree_prediction <- predict(dtree_modfit, test_data, type = "class")

## get error estimates via confusion matrix

 
confusionMatrix(dtree_prediction, as.factor(test_data$classe))


# Random Forest

# fit model to train_data

rf_modfit <- randomForest(as.factor(classe) ~., data = train_data)

# use model on test_data

rf_prediction <-  predict(rf_modfit, test_data, type = "class")

# estimate error via confusion matrix

confusionMatrix(rf_prediction, as.factor(test_data$classe))


```

```{r, quiz}

## test the model to predict 20 different test cases

predict_sub <- predict(rf_modfit, test_set, type = "class")
#predict_sub  ## hidden on purpose


```

 
